---
title: "Creating graphs in R with `ggplot2`"
author: "Thomas Bao"
date: "2023-07-05"
output: html_document
---

# Plots with ggplot2

We are going to make plots in R and it's going to be amazing.

## Loading in necessary packages.

```{r load-in}
library(tidyverse)
library(readxl)
library(here)
```

## Loading in data via csv and seeing basic stats.
We'll be loading in the ca_np.csv file into an object named "ca_np"

```{r csv import}
ca_np <- read_csv(here("data", "ca_np.csv"))
```

Here are some additional functions to view additional details about the dataset:

To view the the variable/column names -
```{r names}
names(ca_np)
```

To see the first rows of the data (first 6 by default):
```{r head}
head(ca_np)
```

To see a quick summary of the data:
```{r summary}
summary(ca_np)
```

## Loading in data via xlsx.
Use a similar function to read xlsx (Excel) files instead.

```{r readxl -> read_excel}
ci_np <- read_excel(here("data", "ci_np.xlsx"))
```

**If the file type is known, you can use "read_xlsx()" or "read_xls()" as well.

## Making our first ggplot - Visitors to Channel Islands NP
What to tell R
1. We are using the function ggplot()
2. What data we're using and what variables will be plot 
3. The type of graph we are going to plot on (i.e. dot, line, column/bar)

Generic form:
ggplot(data = df_name, aes(x = x_var_name, y = y_var_name)) + geom_type()

Here is a plot of visitors to Channel Islands NP over time:
```{r plot Channel Islands visitors over years}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()
```

We're going to set it as object named "geom_base" so we can modify/add-on.
```{r plot Channel Islands visitors over years -> geom_base}
geom_base <- ggplot(data = ci_np, aes(x = year, y = visitors))
```

We can change the plot type by using + geom_type().
```{r change plot type to point}
geom_base +
  geom_point()
```
```{r change plot type to columns}
geom_base +
  geom_col()
```
```{r change plot type to area}
geom_base +
  geom_area()
```

## Customizing ggplot plots

### Aesthetics
Aesthetics can be modified by adding arguments to the geom_type() function. These include:
- color = or colour = : update point or line colors

- fill = : update fill color for objects with areas

- linetype = : update the line type (dashed, long dash, etc.)

- pch = : update the point style

- size = : update the element size (e.g. of points or line thickness)

- alpha = : update element opacity (1 = opaque, 0 = transparent)

Let's make our plot a line plot with purple, dashed lines.
```{r geom_base with purple, dashed lines}
geom_base +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )
```
### Mapping variables onto aesthetics
Variables can also be mapped onto aesthetics. For example, changing the shading of a point depending on a different variable.
These variables should be placed inside the aes() function inside of a geom_type() function:
- geom_type(aes(size = var_name, etc.))

```{r Example of mapping variables onto aesthetics}
geom_base +
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.6
  )
```